Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2019-05-01 SupportingInformation.Rmd
typo fixed and plot label modified

Once character evolution was simulated 100 times for each dependence structure, 
we had 817 bivariate characters for each of 500 simulations. We thus analyzed 500 corresponding 
bivariate copulas (one for each simulation), each comprised of 817 points. 
To assess the symmetry or asymmetry of tail dependence for each copula, we computed 
the three non-parametric statistics $\cor_{0,0.2}-\cor{0.8,1}$, $\Ps_{0,0.2}-\Ps{0.8,1}$,
and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ for each copula. For each statistic and for each
dependence structure, we made a histogram (Fig. \ref{fig_cop_evo}). Values of these statistics less than 0
indicate lower-tail dependence in characters of extant species, whereas values greater
than 0 indicate upper-tail dependence. Asymmetry of tail dependence in evolutionary changes to characters
was associated with similar asymmetry of tail dependence in extant characters.

@@

<!--adding plots showing results to confirm hypothesis behind third mechanism-->
\begin{figure}[!h]
\textbf{ \hspace{6.5 cm} (A) \hspace{2 cm} (B) \hspace{2 cm} (C)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeL_hist_taildep.pdf}\\
\end{center}
\textbf{ \hspace{6.5 cm} (D) \hspace{2 cm} (E) \hspace{2 cm} (F)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatL_hist_taildep.pdf}\\
\end{center}
\textbf{ \hspace{6.5 cm} (G) \hspace{2 cm} (H) \hspace{2 cm} (I)} \\
\vspace{-0.7 cm}
\begin{center}
\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/BivNormal_hist_taildep.pdf}\\
\end{center}
\textbf{ \hspace{6.5 cm} (J) \hspace{2 cm} (K) \hspace{2 cm} (L)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatR_hist_taildep.pdf}\\
\end{center}
\textbf{ \hspace{6.5 cm} (M) \hspace{2 cm} (N) \hspace{2 cm} (O)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeR_hist_taildep.pdf}\\
\caption[Non-parametric statistics on a mechanism of tail dependence in character distributions]{Three 
measures of asymmetry of tail dependence ($\cor_{0,0.2}-\cor_{0.8,1}$ (A, D, G, J, M), 
$\Ps_{0,0.2}-\Ps_{0.8,1}$ (B, E, H, K, N), and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ (C, F, I, L, O)) between two characters, across extant species,  were 
computed for each of 100 simulations of mammalian evolution for each of five types of tail dependence between 
evolutionary changes in the characters (extreme left-tail dependence (A-C), moderate left-tail dependence (D-F), 
symmetry of tail dependence (G-I), moderate right-tail dependence (J-L), and extreme right-tail dependence 
(M-O)). Asymmetry of tail dependence in evolutionary changes in characters tended to beget similar
asymmetry in the characters across extant species. For additional details see 
Appendix \ref{evol_mech}. The number in the upper right of each panel indicates how many statistics out of $100$ were less than or equal to $0$. Values substantially lower (respectively, higher) than $50$ therefore indicate greater left (respectively, right) tail-dependence between characters.\label{fig_cop_evo}}
\end{center}
\end{figure}

>>>>

Once character evolution was simulated 100 times for each dependence structure, 
we had 817 bivariate characters for each of 500 simulations. We thus analyzed 500 corresponding 
bivariate copulas (one for each simulation), each comprised of 817 points. 
To assess the symmetry or asymmetry of tail dependence for each copula, we computed 
the three non-parametric statistics $\cor_{0,0.2}-\cor_{0.8,1}$, $\Ps_{0,0.2}-\Ps_{0.8,1}$,
and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ for each copula. For each statistic and for each
dependence structure, we made a histogram (Fig. \ref{fig_cop_evo}). Values of these statistics greater than 0
indicate lower-tail dependence in characters of extant species, whereas values less
than 0 indicate upper-tail dependence. Asymmetry of tail dependence in evolutionary changes to characters
was associated with similar asymmetry of tail dependence in extant characters.

@@

<!--adding plots showing results to confirm hypothesis behind third mechanism-->
\begin{figure}[!h]
\textbf{ \hspace{6.5 cm} (A) \hspace{2 cm} (B) \hspace{2 cm} (C)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeL_hist_taildep.pdf}\\
\end{center}
\textbf{ \hspace{6.5 cm} (D) \hspace{2 cm} (E) \hspace{2 cm} (F)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatL_hist_taildep.pdf}\\
\end{center}
\textbf{ \hspace{6.5 cm} (G) \hspace{2 cm} (H) \hspace{2 cm} (I)} \\
\vspace{-0.7 cm}
\begin{center}
\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/BivNormal_hist_taildep.pdf}\\
\end{center}
\textbf{ \hspace{6.5 cm} (J) \hspace{2 cm} (K) \hspace{2 cm} (L)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatR_hist_taildep.pdf}\\
\end{center}
\textbf{ \hspace{6.5 cm} (M) \hspace{2 cm} (N) \hspace{2 cm} (O)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeR_hist_taildep.pdf}\\
\caption[Non-parametric statistics on a mechanism of tail dependence in character distributions]{Three 
measures of asymmetry of tail dependence ($\cor_{0,0.2}-\cor_{0.8,1}$ (A, D, G, J, M), 
$\Ps_{0,0.2}-\Ps_{0.8,1}$ (B, E, H, K, N), and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ (C, F, I, L, O)) between two characters, across extant species,  were 
computed for each of 100 simulations of mammalian evolution for each of five types of tail dependence between 
evolutionary changes in the characters (extreme left-tail dependence (A-C), moderate left-tail dependence (D-F), 
symmetry of tail dependence (G-I), moderate right-tail dependence (J-L), and extreme right-tail dependence 
(M-O)). Asymmetry of tail dependence in evolutionary changes in characters tended to beget similar
asymmetry in the characters across extant species. For additional details see 
Appendix \ref{evol_mech}. The number above each panel indicates how many statistics out of $100$ were less than or equal to $0$. Values substantially lower (respectively, higher) than $50$ therefore indicate greater left (respectively, right) tail-dependence between characters.\label{fig_cop_evo}}
\end{center}
\end{figure}
